### 2019-11-16 main_text/main_text.tex
Another round of edits by MJM

	There are two significant mechanisms for transport of xanthosine across
	the cell membrane. In the induced system, the main transporter is XapB,
	whereas in the uninduced system, there is almost no XapB. Instead,
	xanthosine can enter the cell through the two nucleoside transporters
	NupC and NupG, which have a very low affinity for
	xanthosine~\cite{Norholm2001}. All these transporters, XapB, NupC, and
	NupG, are powered by the proton gradient across the
	membrane~\cite{Norholm2001}, which is why we assume their kinetic scheme
	to be similar to that of the \emph{lac} permease (as it is described
	in~\cite{Kaback2015}). There can be import and export of xanthosine, and
	which one dominates depends on the proton and xanthosine concentrations
	on the two sides of the membrane. In both cases, a proton and a
	substrate need to bind to the transporter on one side of the membrane
	and detach from it on the other side before the empty transporter moves
	back to the other side again. We refer the reader to~\nameref{S1_Text}
	for a detailed description of the transport and its modeling, and just
	state the result here. We model influx and efflux separately. For XapB,
	we use Michaelis-Menten kinetics with parameters $k_{\n{b, i}}$,
	$K_{\n{b, i}}$ for influx and $k_{\n{b, e}}, K_{\n{b,e}}$ for efflux.
	For the Nup transporters, we also use Michaelis-Menten kinetics but can linearize the Michaelis-Menten term $k_{\n{cat}} \frac{\n{[x]}}{K_{\n{M}} + \n{[x]}} \approx \tilde{k} \n{[x]}$ because of the transporter's low affinity for xanthosine (i.e. $K_{\n{M}} \gg \n{[x]}$). For the rate parameters $\tilde{k}$, we write $k_{\n{nup}}$ for
	influx and $\xi k_{\n{nup}}$ for efflux.

@@

	\subsection*{Stochastic simulations}
	Stochastic simulations of the full 3-dimensional system of mRNA, protein,
	and xanthosine were run for comparison with the deterministic results.
	In~\nameref{S1_Text}, we present the underlying chemical master equation of
	the system. Because of the two different fixed points at low and high
	expression, the copy numbers in the problem vary from less than five to several thousand.
	The larger the amounts of mRNA, protein, and xanthosine, the larger the number of reactions, and thus, the larger the computational cost of Gillespie's classical algorithm. This makes Gillespie's \texttau-leap algorithm ideal for the high expression state. On the
	other hand, \texttau-leaping cannot be used for the small copy numbers in the low expression state. For
	these reasons, we chose to work with the algorithm described
	in~\cite{Cao2006}, a hybrid form between Gillespie's classical and his
	\texttau-leap algorithm. We gratefully worked with the Python implementation
	of this algorithm in \emph{StochPy}, version 2.3~\cite{Maarleveld2015}.

@@

	We have found that no bistability can emerge from the genetic circuit unless
	at least one component has two binding sites for its activator. Additional
	binding sites or cooperativity seem to increase the stability of the switch.
	In addition, simply knowing the experimental switching concentration of
	xanthosine permits us, for example, to infer the approximate value of the
	dissociation constant between the transcription factor XapR and the inducer
	xanthosine. The value we found is roughly one to two orders of magnitude larger than what has been measured for LacI and IPTG~\cite{RazoMejia2018}, meaning the interaction of XapR and xanthosine is rather weak.

@@

	\section*{Acknowledgments}
	We thank Griffin Chure, Charlotte Strandkvist, and Kyle Naughton for
	providing their data on the \emph{xap} system and pointing out the
	bistability. We also thank Griffin Chure for answering the many biological
	questions that arose, Nathan Belliveau for bringing similar systems to our
	attention, and Jane Kondev and Jin Wang for interesting discussions.
	For assistance with illustrations, we also thank Nigel Orme and Griffin Chure.
	This work was supported by 1R35 GM118043-01 Maximizing Investigators' Research Award (MIRA) (to R.P.), the National Science Foundation
	Graduate Research Fellowship under Grant No.\ DGE-114446 (M.J.M.). and the
	Werner Siemens Foundation through the Swiss Study Foundation (K.S.L.).

>>>>

	There are two significant mechanisms for transport of xanthosine across
	the cell membrane. In the induced system, the main transporter is XapB,
	whereas in the uninduced system, there is almost no XapB. Instead,
	xanthosine can enter the cell through the two nucleoside transporters
	NupC and NupG, which have a very low affinity for
	xanthosine~\cite{Norholm2001}. All these transporters, XapB, NupC, and
	NupG, are powered by the proton gradient across the
	membrane~\cite{Norholm2001}, which is why we assume their kinetic scheme
	to be similar to that of the \emph{lac} permease (as it is described
	in~\cite{Kaback2015}). There can be import and export of xanthosine, and
	which one dominates depends on the proton and xanthosine concentrations
	on the two sides of the membrane. In both cases, a proton and a
	substrate need to bind to the transporter on one side of the membrane
	and detach from it on the other side before the empty transporter moves
	back to the other side again. We refer the reader to~\nameref{S1_Text}
	for a detailed description of the transport and its modeling, and just
	state the result here. We model influx and efflux separately. For XapB,
	we use Michaelis-Menten kinetics with parameters $k_{\n{b, i}}$,
	$K_{\n{b, i}}$ for influx and $k_{\n{b, e}}, K_{\n{b,e}}$ for efflux.
	For the Nup transporters, we also use Michaelis-Menten kinetics but,
	because of the transporter's low affinity for xanthosine,
	we can linearize the Michaelis-Menten term as
	$k_{\n{cat}} {\n{[x]}}/(K_{\n{M}} + \n{[x]}) \approx \tilde{k} \n{[x]}$
	(i.e., $K_{\n{M}}~\gg~\n{[x]}$ across the physiologically relevant range for $\n{[x]}$).
	For the rate parameters $\tilde{k}$, we write $k_{\n{nup}}$ for
	influx and $\xi k_{\n{nup}}$ for efflux.

@@

	\subsection*{Stochastic simulations}
	Stochastic simulations of the full 3-dimensional system of mRNA, protein,
	and xanthosine were run for comparison with the deterministic results.
	In~\nameref{S1_Text}, we present the underlying chemical master equation of
	the system. Because of the two different fixed points at low and high
	expression, the protein copy numbers in the problem vary from less than
	five to several thousand. Even worse, xanthosine copy numbers may range
	as high as $10^7$ at the high expression fixed point. For such large
	copy numbers, the number of reaction firings that must be simulated with
	Gillespie's classical algorithm leads to an impractical computational
	cost. This would make Gillespie's \texttau-leap algorithm ideal for the high
	expression state. On the other hand, \texttau-leaping cannot be used for
	the small protein copy numbers in the low expression state, or the mRNA
	copy number which remains of order ten or less in both states. For
	these reasons, we chose to work with the algorithm described
	in~\cite{Cao2006}, a hybrid form between Gillespie's classical and his
	\texttau-leap algorithm. We gratefully worked with the Python implementation
	of this algorithm in \emph{StochPy}, version 2.3~\cite{Maarleveld2015}.

@@

	We have found that no bistability can emerge from the genetic circuit unless
	at least one component has two binding sites for its activator. Additional
	binding sites or cooperativity seem to increase the stability of the switch.
	In addition, simply knowing the experimental switching concentration of
	xanthosine permits us, for example, to infer the approximate value of the
	dissociation constant between the transcription factor XapR and the inducer
	xanthosine. The value we infer is roughly one to two orders of magnitude
	larger than what has been measured for LacI and
	IPTG~\cite{RazoMejia2018}, meaning the interaction of XapR and
	xanthosine is rather weak.

@@

	\section*{Acknowledgments}
	We thank Griffin Chure, Charlotte Strandkvist, and Kyle Naughton for
	providing their data on the \emph{xap} system and pointing out the
	bistability. We also thank Griffin Chure for answering the many biological
	questions that arose, Nathan Belliveau for bringing similar systems to our
	attention, and Jane Kondev and Jin Wang for interesting discussions.
	For assistance with illustrations, we also thank Nigel Orme and Griffin Chure.
	This work was supported by 1R35 GM118043-01 Maximizing Investigators'
	Research Award (MIRA) (to R.P.), and the Werner Siemens Foundation
	through the Swiss Study Foundation (K.S.L.). This material is based upon
	work supported by the National Science Foundation Graduate Research
	Fellowship under Grant No.\ DGE-1745301 (M.J.M.).
